Something about Continuous Deductive Reasoning over Streamed RDF Data, maybe claim Soundness or some other Theoretical Quality

### 2015-06-26 content/contribution/continuous-datalog.tex
Converted to axiom environment

	In regards to Continuous Datalog, an ``instant'', a ``stream of facts'' and a ``sliding window'' over a stream are defined as follows:
	\begin{description}
		\item[Instant\label{def:continuous datalog: instant}]\hfill\\
			An instant is a discrete point in time.
			The continuous distance in time between these discrete points may be theoretically infinitesimal, but will in practise have a lower bound set by hardware limitations, or will be defined as part of some specific implementation.
			In these cases, anything that occurs between discrete instants is considered to have occurred at the next instant.
		\item[Stream of Facts\label{def:continuous datalog: stream}]\hfill\\
			A stream of facts is a \emph{sequence} of Datalog facts ordered by some temporal characteristic of each instance of each fact, e.g. the time at which each was entailment by some process at the source of the stream.
			Each instance of a fact is uniquely identifiable by the combination of its signature (its predicate $p$, subject $s$ and object $o$) and its entailment time $t$, so an instance is denoted ${p(s,o):t}$.
			This distinction between instances of facts within a stream is represented in Axiom~\ref{axiom:continuous datalog: window disjointness}.
		\item[Sliding Window\label{def:continuous datalog: window}]\hfill\\
			A sliding window of range ${X \in \mathbb{N}}$ over any stream $S$ at any given instant ${i \in \mathbb{N}}$ is the \emph{set} of instances of facts that occur in $S$ during the interval ${(i-X,i]}$, and is denoted ${W^{S,X}_{i}}$.
			The concept of a sliding window within the semantics of Continuous Datalog are further defined by Axioms~\ref{axiom:continuous datalog: window range leq 0} and~\ref{axiom:continuous datalog: window composition}.
	\end{description}
	The following axioms define the semantics of streams and windows over those streams in Continuous Datalog.
	\begin{enumerate}\newcounter{continuousDatalogAxioms}
		\item\label{axiom:continuous datalog: window range leq 0}
			A window of range ${X \leq 0}$ over any stream $S$ at any instant ${i \in \mathbb{N}}$ is always empty.
			\begin{equation*}
				W^{S,X}_{i} = \varnothing
			\end{equation*}
		\item\label{axiom:continuous datalog: window composition}
			A window of any range ${X \in \mathbb{N}}$ over any stream $S$ at any given instant ${i \in \mathbb{N}}$ is equivalent to the union of the $X$ consecutive windows of range 1 over the stream $S$ preceding and including that at instant $i$.
			\begin{equation*}
				W^{S,X}_{i} = \mathop{\cup}_{i=0}^{X-1} W^{S,1}_{i-i}
			\end{equation*}
		\item\label{axiom:continuous datalog: window disjointness}
			A window of any range ${X \in \mathbb{N}}$ over any stream $S$ at any instant ${i \in \mathbb{N}}$ is disjoint from a second window of any range ${Y \in \mathbb{N}}$ over $S$ at any instant after ${i - X}$.
			\begin{equation*}
				\forall X' \geq X \, \left( W^{S,X}_{i} \cap W^{S,Y}_{i-X'} = \varnothing \right)
			\end{equation*}
		% \item\label{axiom:continuous datalog: stream disjointness}
		% 	The contents of any two streams are disjoint, thus a window of any range ${X \in \mathbb{N}}$ over any stream $S$ at any instant ${i \in \mathbb{N}}$ is disjoint from a second window of any range ${Y \in \mathbb{N}}$ over any other stream ${Z \neq S}$ at any instant ${i \in \mathbb{N}}$.
		% 	\begin{equation*}
		% 		W^{S,X}_{i} \cap W^{Z,Y}_{i} = \varnothing
		% 	\end{equation*}
		\setcounter{continuousDatalogAxioms}{\theenumi}
	\end{enumerate}

@@

	From this description of \emph{instance entailment}, Axiom~\ref{axiom:continuous datalog: entailment disjointness} may be derived.
	\begin{enumerate}\setcounter{enumi}{\thecontinuousDatalogAxioms}
		\item\label{axiom:continuous datalog: entailment disjointness}
			If two sets, $A_{inst}$ and $B_{inst}$, of axiom instances are disjoint, then the any set $E_{inst}$ of axiom instances each justified at least in part by $A_{inst}$ is necessarily disjoint from any set $F_{inst}$ of axiom instances each justified at least in part by $B_{inst}$, because entailed fact instances are distinct with regards to the fact instances that justify them.
			\begin{multline*}
				\left( A_{inst} \cap B_{inst} = \varnothing \right) \wedge \left( A_{inst} \vDash E_{inst} \right) \\
					\wedge \left( B_{inst} \vDash F_{inst} \right) \rightarrow \left( E_{inst} \cap F_{inst} = \varnothing \right)
			\end{multline*}
		\setcounter{continuousDatalogAxioms}{\theenumi}
	\end{enumerate}

	Given these definitions, a Continuous Datalog program may be formally defined, both in terms of streams and in terms of its instantaneous states:
	\begin{description}
		\item[Continuous Datalog Program\label{def:continuous datalog: CDP}]\hfill\\
			A Continuous Datalog Program (${CDP}$) is composed of a set of Positive Datalog axioms $A$, a set of streams of facts $S$ and the range ${R \in \mathbb{N}}$ of the windows over each stream in $S$.
			This specifies that the streamed data from $S$ is to be considered valid from its arrival for the duration represented by $R$.
			Such a continuous program entails a triple composed of the traditional entailments of ${A \vDash A'}$, a stream of entailed instances $P$ and a stream of instance negations $N$, whose contents are defined in Axioms~\ref{axiom:continuous datalog: positive window increment},~\ref{axiom:continuous datalog: negative window increment} and~\ref{axiom:continuous datalog: entailment precedes negation}.
			Streams $P$ and $N$ are therefore the continuous reasoning equivalents of the streams produced by the \emph{IStream} and \emph{DStream} operators of the \emph{CQL} continuous query language \citep{CQL} respectively.
			The truth value of a ${CDP}$ is instantaneous, with the the truth function of Continuous Datalog $T^{CD}$ taking both a program ${CDP}$ and an instant ${i \in \mathbb{N}}$ as arguments.
			\begin{equation*}
				CDP = \{A,S,R\} \vDash \{A',P,N\}
			\end{equation*}
		\item[State of a ${CDP}$\label{def:continuous datalog: CDPt}]\hfill\\
			The state of a ${CDP}$ at a given instant ${i \in \mathbb{N}}$ (${CDP_{i}}$) is the union of the axioms $A$ and the window $W^{S,R}_{i}$.
			As both $A$ and ${trad\left( W^{S,R}_{i} \right)}$ are sets of Datalog axioms, a ${CDP_{i}}$ constitutes a valid Datalog program that may be evaluated by the truth function of Datalog $T^{D}$, the result of which is the definition for the instantaneous truth of the ${CDP}$ at instant $i$ by $T^{CD}$.
			The instantaneous entailments of a $CDP_{i}$ are the union of the persistent entailments of ${A \vDash A'}$ (which are guaranteed to be entailed by every state, as Positive Datalog is monotonic) and the set $E_{i}$ of instances of the transient entailments that are justified at least in part by some subset of $W^{S,R}_{i}$. 
			\begin{equation*}
				T^{CD} \left( CDP, i \right) = T^{D} \left( CDP_{i} \right)
			\end{equation*}
			\begin{equation*}
				CDP_{i} = A \cup W^{S,R}_{i} \vDash A' \cup E_{i}
			\end{equation*}
	\end{description}
	The following axioms define the semantics of the streams of entailments $P$ and and negations $N$ of a ${CDP}$ in terms of the instantaneous entailments of the sequence of states of that continuous program.
	\begin{enumerate}\setcounter{enumi}{\thecontinuousDatalogAxioms}
		\item\label{axiom:continuous datalog: positive window increment}
			The window of range 1 at any given instant ${i \in \mathbb{N}}$ over the stream $P$ entailed by a ${CDP}$ is defined as the set of facts that are entailed by the state of the ${CDP}$ at instant $i$ that were not entailed by the state of ${CDP}$ at the previous instant instance ${i-1}$.
			\begin{equation*}
				W^{P,1}_{i} = \left( A' \cup E_{i} \right) \setminus \left( A' \cup E_{i-1} \right) = E_{i} \setminus E_{i-1}
			\end{equation*}
		\item\label{axiom:continuous datalog: negative window increment}
			The window of range 1 at any given instant ${i \in \mathbb{N}}$ over the stream $N$ entailed by a ${CDP}$ is defined as the set of facts that were entailed by the state of the ${CDP}$ at the previous instant instance ${i-1}$ that are no longer entailed by the state of ${CDP}$ at instant $i$.
			\begin{equation*}
				W^{N,1}_{i} = \left( A' \cup E_{i-1} \right) \setminus \left( A' \cup E_{i} \right) = E_{i-1} \setminus E_{i}
			\end{equation*}
		\item\label{axiom:continuous datalog: entailment precedes negation}
			Facts may only be negated by some part of their justification leaving the sliding window $W^{S,R}_{i}$, this necessarily being after they were entailed.
			As such, a window of any range ${X \in \mathbb{N}}$ over the stream of entailments $P$ at any given instant ${i \in \mathbb{N}}$ is disjoint from a window of any range ${Y \in \mathbb{N}}$ over the the stream of entailment negations $N$ at any instant after ${i - X}$
			\begin{equation*}
				\forall X' \geq X \left( W^{P,X}_{i} \cap W^{N,Y}_{i-X'} = \varnothing \right)
			\end{equation*}
		\setcounter{continuousDatalogAxioms}{\theenumi}
	\end{enumerate}

>>>>

	In regards to Continuous Datalog, an ``instant'', a ``stream of facts'' and a ``sliding window'' over a stream are defined as follows:
	\begin{description}
		\item[Instant\label{def:continuous datalog: instant}]\hfill\\
			An instant is a discrete point in time.
			The continuous distance in time between these discrete points may be theoretically infinitesimal, but will in practise have a lower bound set by hardware limitations, or will be defined as part of some specific implementation.
			In these cases, anything that occurs between discrete instants is considered to have occurred at the next instant.
		\item[Stream of Facts\label{def:continuous datalog: stream}]\hfill\\
			A stream of facts is a \emph{sequence} of Datalog facts ordered by some temporal characteristic of each instance of each fact, e.g. the time at which each was entailment by some process at the source of the stream.
			Each instance of a fact is uniquely identifiable by the combination of its signature (its predicate $p$, subject $s$ and object $o$) and its entailment time $t$, so an instance is denoted ${p(s,o):t}$.
			This distinction between instances of facts within a stream is represented in Axiom~\ref{axiom:continuous datalog: window disjointness}.
		\item[Sliding Window\label{def:continuous datalog: window}]\hfill\\
			A sliding window of range ${X \in \mathbb{N}}$ over any stream $S$ at any given instant ${i \in \mathbb{N}}$ is the \emph{set} of instances of facts that occur in $S$ during the interval ${(i-X,i]}$, and is denoted ${W^{S,X}_{i}}$.
			The concept of a sliding window within the semantics of Continuous Datalog are further defined by Axioms~\ref{axiom:continuous datalog: window range leq 0} and~\ref{axiom:continuous datalog: window composition}.
	\end{description}

The following axioms define the semantics of streams and windows over
those streams in Continuous Datalog:

\begin{axiom}
\label{axiom:continuous datalog: window range leq 0}
A window of range ${X \leq 0}$ over any stream $S$ at any instant ${i
  \in \mathbb{N}}$ is always empty.
\begin{equation*}
W^{S,X}_{i} = \varnothing
\end{equation*}
\end{axiom}

\begin{axiom}
\label{axiom:continuous datalog: window composition}
A window of any range ${X \in \mathbb{N}}$ over any stream $S$ at any
given instant ${i \in \mathbb{N}}$ is equivalent to the union of the
$X$ consecutive windows of range 1 over the stream $S$ preceding and
including that at instant $i$.
\begin{equation*}
W^{S,X}_{i} = \mathop{\cup}_{i=0}^{X-1} W^{S,1}_{i-i}
\end{equation*}
\end{axiom}

\begin{axiom}
\item\label{axiom:continuous datalog: window disjointness}
A window of any range ${X \in \mathbb{N}}$ over any stream $S$ at any
instant ${i \in \mathbb{N}}$ is disjoint from a second window of any
range ${Y \in \mathbb{N}}$ over $S$ at any instant after ${i - X}$.
\begin{equation*}
\forall X' \geq X \, \left( W^{S,X}_{i} \cap W^{S,Y}_{i-X'} = \varnothing \right)
\end{equation*}
\end{axiom}

@@

	From this description of \emph{instance entailment}, Axiom~\ref{axiom:continuous datalog: entailment disjointness} may be derived.

\begin{axiom}
\label{axiom:continuous datalog: entailment disjointness}
If two sets, $A_{inst}$ and $B_{inst}$, of axiom instances are
disjoint, then the any set $E_{inst}$ of axiom instances each
justified at least in part by $A_{inst}$ is necessarily disjoint from
any set $F_{inst}$ of axiom instances each justified at least in part
by $B_{inst}$, because entailed fact instances are distinct with
regards to the fact instances that justify them.
\begin{multline*}
\left( A_{inst} \cap B_{inst} = \varnothing \right) \wedge \left( A_{inst} \vDash E_{inst} \right) \\
\wedge \left( B_{inst} \vDash F_{inst} \right) \rightarrow \left( E_{inst} \cap F_{inst} = \varnothing \right)
\end{multline*}
\end{axiom}

	Given these definitions, a Continuous Datalog program may be formally defined, both in terms of streams and in terms of its instantaneous states:
	\begin{description}
		\item[Continuous Datalog Program\label{def:continuous datalog: CDP}]\hfill\\
			A Continuous Datalog Program (${CDP}$) is composed of a set of Positive Datalog axioms $A$, a set of streams of facts $S$ and the range ${R \in \mathbb{N}}$ of the windows over each stream in $S$.
			This specifies that the streamed data from $S$ is to be considered valid from its arrival for the duration represented by $R$.
			Such a continuous program entails a triple composed of the traditional entailments of ${A \vDash A'}$, a stream of entailed instances $P$ and a stream of instance negations $N$, whose contents are defined in Axioms~\ref{axiom:continuous datalog: positive window increment},~\ref{axiom:continuous datalog: negative window increment} and~\ref{axiom:continuous datalog: entailment precedes negation}.
			Streams $P$ and $N$ are therefore the continuous reasoning equivalents of the streams produced by the \emph{IStream} and \emph{DStream} operators of the \emph{CQL} continuous query language \citep{CQL} respectively.
			The truth value of a ${CDP}$ is instantaneous, with the the truth function of Continuous Datalog $T^{CD}$ taking both a program ${CDP}$ and an instant ${i \in \mathbb{N}}$ as arguments.
			\begin{equation*}
				CDP = \{A,S,R\} \vDash \{A',P,N\}
			\end{equation*}
		\item[State of a ${CDP}$\label{def:continuous datalog: CDPt}]\hfill\\
			The state of a ${CDP}$ at a given instant ${i \in \mathbb{N}}$ (${CDP_{i}}$) is the union of the axioms $A$ and the window $W^{S,R}_{i}$.
			As both $A$ and ${trad\left( W^{S,R}_{i} \right)}$ are sets of Datalog axioms, a ${CDP_{i}}$ constitutes a valid Datalog program that may be evaluated by the truth function of Datalog $T^{D}$, the result of which is the definition for the instantaneous truth of the ${CDP}$ at instant $i$ by $T^{CD}$.
			The instantaneous entailments of a $CDP_{i}$ are the union of the persistent entailments of ${A \vDash A'}$ (which are guaranteed to be entailed by every state, as Positive Datalog is monotonic) and the set $E_{i}$ of instances of the transient entailments that are justified at least in part by some subset of $W^{S,R}_{i}$. 
			\begin{equation*}
				T^{CD} \left( CDP, i \right) = T^{D} \left( CDP_{i} \right)
			\end{equation*}
			\begin{equation*}
				CDP_{i} = A \cup W^{S,R}_{i} \vDash A' \cup E_{i}
			\end{equation*}
	\end{description}
	The following axioms define the semantics of the streams of entailments $P$ and and negations $N$ of a ${CDP}$ in terms of the instantaneous entailments of the sequence of states of that continuous program.

\begin{axiom}
\label{axiom:continuous datalog: positive window increment}
The window of range 1 at any given instant ${i \in \mathbb{N}}$ over
the stream $P$ entailed by a ${CDP}$ is defined as the set of facts
that are entailed by the state of the ${CDP}$ at instant $i$ that were
not entailed by the state of ${CDP}$ at the previous instant instance
${i-1}$.
			\begin{equation*}
				W^{P,1}_{i} = \left( A' \cup E_{i} \right) \setminus \left( A' \cup E_{i-1} \right) = E_{i} \setminus E_{i-1}
			\end{equation*}
\end{axiom}

\begin{axiom}
\label{axiom:continuous datalog: negative window increment}
			The window of range 1 at any given instant ${i \in \mathbb{N}}$ over the stream $N$ entailed by a ${CDP}$ is defined as the set of facts that were entailed by the state of the ${CDP}$ at the previous instant instance ${i-1}$ that are no longer entailed by the state of ${CDP}$ at instant $i$.
			\begin{equation*}
				W^{N,1}_{i} = \left( A' \cup E_{i-1} \right) \setminus \left( A' \cup E_{i} \right) = E_{i-1} \setminus E_{i}
			\end{equation*}
\end{axiom}

\begin{axiom}
\label{axiom:continuous datalog: entailment precedes negation}
			Facts may only be negated by some part of their justification leaving the sliding window $W^{S,R}_{i}$, this necessarily being after they were entailed.
			As such, a window of any range ${X \in \mathbb{N}}$ over the stream of entailments $P$ at any given instant ${i \in \mathbb{N}}$ is disjoint from a window of any range ${Y \in \mathbb{N}}$ over the the stream of entailment negations $N$ at any instant after ${i - X}$
			\begin{equation*}
				\forall X' \geq X \left( W^{P,X}_{i} \cap W^{N,Y}_{i-X'} = \varnothing \right)
			\end{equation*}
\end{axiom}
